---
title: "Lab 4 - Writing Functions II"
author: "Cameron Kay & Ashley Miller"
date: "5/22/2019"
output: 
  html_document:
    theme: kate
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
# load packages
library(tidyverse)
```

There are two parts to this lab. The first includes writing a relatively complicated function. The second involves using that function within a relatively basic shiny app.

## Part A: Writing a summary function

Write a function that takes a given data frame as its first argument, a grouping variable as its second argument, and a column to summarize as its third argument.

### Function behavior

By default, the function should report:

Total number of observations
* Number of non-missing observations
* Number of missing observations
* Mean
* Standard deviation
* Minimum
* Maximum

However, the function should be flexible so it can take any function(s) as its input to summarize by, provided the function results in a single value (e.g., anything you could feed to `dplyr::summarize`). You’ll need a fourth optional argument that lists the function(s) to use in the summary.

### Errors

Embed an error if

* Anything other than a data frame is supplied as the first argument
* The provided column to summarize is not numeric.

### Warnings

Embed a warning if

* The grouping variable is numeric (generally you would expect * the grouping variable to be a categorical variable)

### Some guidance

If you want to use {dplyr} you’ll need to use non-standard evaluation (NSE), so something like `!!enquo(group_var)`, where `group_var` is the formal argument of your function. If you go this route, you will also likely want to look into `summarize_*`, which take an additional .funs argument (to make the functions used in summarizing more flexible). Finally, dplyr::pull (together with NSE) will likely be helpful for checking types for warnings/errors.

Alternatively, you might consider `split`ing the dataset by the grouping variable, then looping over each split to apply the function(s) supplied.

### Sample output

See the [assignment page](https://rmarkdown.rstudio.com/authoring_basics.html) for the examples.

```{r parta}
# create the summariser funcntion
summariser <- function(data, group, column, FUN, na.rm = FALSE) {
  # return error if data is not of type data
  if (!is.data.frame(data)) {
    stop("Data is not of type data.frame")
  }
  # return error if the provided columnt to summarize is not of type numeric
  if (!is.numeric(data[, column])) {
    stop("The column to be summarized is not of type numeric.")
  }
  # return warning if the grouping variable is numeric
  if (is.numeric(data[, group])) {
    warning("The grouping variable was of type numeric and was coerced to type factor.")
  }
  # if function is missing, use default functions
  if (missing(FUN)) {
    FUN <- list(n       = length,
                n_valid = function(x) sum(!is.na(x)),
                n_miss  = function(x) sum(is.na(x)),
                m       = function(x) mean(x, na.rm = na.rm),
                sd      = function(x) sd(x, na.rm = na.rm),
                min     = function(x) min(x, na.rm = na.rm),
                max     = function(x) max(x, na.rm = na.rm))
  }
  # split the data by the group column
  data_split <- split(data, data[ , group])
  # summarise the column of interest
  data_summed <- lapply(data_split, function(x){
    vapply(FUN, FUN.VALUE = numeric(1), function(y) {
      y(x[, column])
    })
  })
  # rbind list
  data_summed  <- do.call(rbind, data_summed)
  # apply names to columns
  colnames(data_summed) <- names(FUN) 
  # return result
  data_summed
}

# testing

summariser(iris, "Species", "Sepal.Length") # works

summariser("data", "Species", "Sepal.Length") # throw data.frame error

summariser(mtcars, "cyl", "wt") # throws coercing warning

summariser(airquality, "Month", "Ozone") # returns NAS because na.rm = FALSE

summariser(airquality, "Month", "Ozone", na.rm = TRUE) # does not return NAS because na.rm = TRUE

summariser(iris, "Species", "Sepal.Length", FUN = list(mean     = mean,
                                                       sd       = sd,
                                                       root_sum = function(x) sum(sqrt(x)))) # works with custom functions

```

## Part B: Shiny app

Build a shiny app using the `gss_cat` dataset available through the {forcats} package, which you can access with `data(gss_cat, package = "forcats")`.

The app should allow the user to select a variable, including (potentially) “none”, and will produce a histogram of `tvhours` faceted by that variable. Below the plot, render a table summarizing the distributions using the function you wrote in Part A. Note, I would encourage you to take a look at `selectInput` rather than `radioButtons` for the categorical variable, but either should work.

For an example of what the shiny app should look like, see [here](https://djanderson07.shinyapps.io/lab-c3/).
